---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# exactplot

<!-- badges: start -->
<!-- badges: end -->

The goal of `exactplot` is to produce millimeter-exact figure layouts and annotate your plots with the full power of Latex. `exactplot` wraps around [`tikzDevice`](https://daqana.github.io/tikzDevice/) and provides utility functions to compose grid plots (e.g., `ggplot2` output, but not base plots).

## Disclaimer

This is a *work-in-progress* package and mainly serves as a collection of scripts that I have used in my past papers to produce my scientific figures. I currently do not plan to release it on CRAN and I am only planning to add features if I need them for my work. If you would like to see some feature included, please open a pull request, or feel free to fork the package.

## Installation

You can install the development version of `exactplot` from Github:

``` r
devtools::install_github("const-ae/exactplot")
```

You might need to install additional fonts. By default, `exactplot` uses the [IBM Plex](https://github.com/IBM/plex) font family. On Mac, you can install them using brew.

``` shell
brew install --cask font-ibm-plex-sans font-ibm-plex-mono font-ibm-plex-math
```

## Example

```{r, eval=FALSE}
library(exactplot)
library(tidyverse)
library(palmerpenguins)
```
```{r, include=FALSE}
devtools::load_all(".")
library(tidyverse)
library(palmerpenguins)
```

The `xp_init` function sets consistent fonts and font sizes, adds additional packages to the `options("tikzLatexPackages")`, and sets the default `ggplot2` theme. The default font sizes are available through the `xp` object. This is needed when you, want to use larger or smaller text in `geom_text()` (remember to change the `size.unit = "pt"`).

```{r}
xp_init()
xp$fontsize
```


Make some simple example plots.

```{r, paged.print=FALSE}
scatter_plot <- penguins |>
  mutate(label = paste0(species, " (", island, ", $", year,"$)")) |>
  ggplot(aes(x = flipper_length_mm, y = bill_length_mm)) +
    geom_point(aes(color = species), show.legend = FALSE) +
    geom_label(data = \(x) slice_max(x, bill_length_mm, by = species), 
               aes(label = label, hjust = ifelse(flipper_length_mm < 200, 0, 1)), vjust = -0.2) +
    labs(x = "Flipper length in $10^{-3}$ m", y = "Bill length in $10^{-3}$ m") +
    coord_cartesian(clip = "off")

beak_lengths <- penguins |>
  ggplot(aes(x = bill_length_mm)) +
   geom_density(aes(fill = species), alpha = 0.7 ) +
   labs(x = "Bill length in $10^{-3}$ m")
```


Combine the two plots with some additional annotations. If you call this function without specifying the filename, you get a quick preview without the latex rendering. By default, `exactplot` uses LuaLatex for rendering, because of its superior font support.

```{r}
xp_compose_plots(
  xp_text("Welcome to \\texttt{exactplot} with \\LaTeX{} support", x = 1, y = 2, fontsize = xp$fontsize_large, fontface = "bold"),
  xp_text("A) Flipper length vs.\\ beak size", x = 1, y = 8),
  xp_plot(scatter_plot, x = 0, y = 12, width = 80, height = 40),
  
  xp_text("B) Beak sizes", x = 84, y = 8),
  xp_plot(beak_lengths, x = 82, y = 12, width = 58, height = 40),
  xp_text("$\\textrm{density}(x) = \\frac{1}{n\\sigma}\\frac{1}{\\sqrt{2\\pi}}\\sum_{i=1}^N{\\exp\\left(\\frac{-(x-x_i)^2}{2\\sigma^2}\\right)}$",
           x = 99, y = 12, fontsize = xp$fontsize_small),
  
  width = 140, height = 52,
  keep_tex_file = FALSE, filename = "man/example.pdf"
)
```


Display the output:

```{r}
print(magick::image_read_pdf("man/example.pdf", density = 600), info = FALSE)
```


